Software Citation Principles

### 2016-03-02 sc-principles/software-citation-principles.tex

\rowcolors{5}{white}{gray!25}
\begin{table}[htbp]
\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases} \flnote{Also, I noticed that some entries have the same requirements: maybe they could be combined to one line?} \dmnote{I would leve them separate. What about adding a column for license?}}
\centering
\scriptsize\setlength{\tabcolsep}{2.5pt}
\begin{tabular}{@{}l l c c c c c c c c@{}}
\toprule
 & & \multicolumn{6}{c}{Requirements} \\
 \cmidrule{3-10}
\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{UID} & Indexed & \multirow{2}{*}{Role} \\
& & name &  &  &  date & repository &  & citations & \\
\midrule
Researcher            & use someone else's software for a paper      & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Researcher            & use someone else's software for new software & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Researcher            & contribute to software                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet \\
Researcher            & find citations of software               & \textbullet &             &             &             &             & \textbullet & \textbullet &             \\
Researcher            & get credit for software development          & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
Researcher            & ``reproduce'' analysis                   & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Researcher            & benchmark software                       & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Researcher            & find software to implement task          & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             \\
Publisher             & publish software paper                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Publisher             & publish papers that cite software        & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             \\
Indexer               & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             \\
Domain group          & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Library\slash archive & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Repository            & show scientific impact of holdings       & \textbullet &             &             &             &             & \textbullet & \textbullet &             \\
Funder                & show how funded software has been used   & \textbullet &             &             &             &             & \textbullet & \textbullet &             \\
\bottomrule
\end{tabular}
\label{tab:use_cases}
\end{table}%

@@

\begin{enumerate}
\item \textbf{Importance}:
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data. \katznote{say specifically that this should be part of the bibliography of a paper?}
\item \textbf{Credit and Attribution}
\begin{enumerate}
\item Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
\item In scholarly literature and other products, whenever and wherever a claim relies upon software, the corresponding software should be cited.
\end{enumerate}
\item \textbf{Unique Identification}
A software citation should include a persistent method for identification that is machine actionable, globally unique, and widely used by a community.
\item \textbf{Access}
Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.
\item \textbf{Persistence}
Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
\item \textbf{Specificity and Verifiability}
Software citations should facilitate identification of, access to, and verification of the specific software that support a claim. Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific version and/or portion of software retrieved subsequently is the same as was originally cited.
\item \textbf{Interoperability and Flexibility}
Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}

>>>>

\rowcolors{5}{white}{gray!25}
\begin{table}[htbp]
\caption{Use cases for software citation, adapted from \cite{SC-Use-Cases} \flnote{Also, I noticed that some entries have the same requirements: maybe they could be combined to one line?} \dmnote{I would leave them separate. What about adding a column for license?}}
\centering
\scriptsize\setlength{\tabcolsep}{2.5pt}
\begin{tabular}{@{}l l c c c c c c c c@{}}
\toprule
 & & \multicolumn{6}{c}{Requirements} \\
 \cmidrule{3-10}
\multirow{2}{*}{Stakeholder} &	\multirow{2}{*}{Use\slash wants to} 	 &  Software  & \multirow{2}{*}{Author(s)} & \multirow{2}{*}{Version \#} & Release & Location\slash  & \multirow{2}{*}{UID} & Indexed & \multirow{2}{*}{Role} \\
& & name &  &  &  date & repository &  & citations & \\
\midrule
Researcher            & use someone else's software for a paper      & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Researcher            & use someone else's software for new software & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Researcher            & contribute to software                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             & \textbullet \\
Researcher            & find citations of software               & \textbullet &             &             &             &             & \textbullet & \textbullet &             \\
Researcher            & get credit for software development          & \textbullet & \textbullet &             & \textbullet & \textbullet & \textbullet &             & \textbullet \\
Researcher            & ``reproduce'' analysis                   & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Researcher            & benchmark software                       & \textbullet &             & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Researcher            & find software to implement task          & \textbullet & \textbullet &             &             & \textbullet & \textbullet & \textbullet &             \\
Publisher             & publish software paper                   & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Publisher             & publish papers that cite software        & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             \\
Indexer               & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             \\
Domain group          & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Library\slash archive & build catalog of software                & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet & \textbullet &             &             \\
Repository            & show scientific impact of holdings       & \textbullet &             &             &             &             & \textbullet & \textbullet &             \\
Funder                & show how funded software has been used   & \textbullet &             &             &             &             & \textbullet & \textbullet &             \\
\bottomrule
\end{tabular}
\label{tab:use_cases}
\end{table}%

@@

\begin{enumerate}
\item \textbf{Importance}:
Software should be considered a legitimate, citable product of research. Software citations should be accorded the same importance in the scholarly record as citations of other research objects, such as publications and data. Specifically, software citations should be part of the reference list for a research product such as a paper.
\item \textbf{Credit and Attribution}
\begin{enumerate}
\item Software citations should facilitate giving scholarly credit and normative and legal attribution to all contributors to the software, recognizing that a single style or mechanism of attribution may not be applicable to all software.\scnote{At least four styles (AMS, APA, Chicago, MLA) for software citation are illustrated with examples in ~\cite{lipson2011cite}.}
\item In scholarly literature and other products, whenever and wherever a claim relies upon software, the corresponding software should be cited.
\end{enumerate}
\item \textbf{Unique Identification}
A software citation should include a persistent method for identification that is machine actionable, globally unique, and widely used by a community.
\item \textbf{Access}
Software citations should facilitate access to the software themselves and to such associated metadata, documentation, data, and other materials, as are necessary for both humans and machines to make informed use of the referenced software.
\item \textbf{Persistence}
Unique identifiers, and metadata describing the software, and its disposition, should persist -- even beyond the lifespan of the software they describe.
\item \textbf{Specificity and Verifiability}
Software citations should facilitate identification of, access to, and verification of the specific software that support a claim. Citations or citation metadata should include information about provenance and fixity sufficient to facilitate verifying that the specific version and/or portion of software retrieved subsequently is the same as was originally cited.
\item \textbf{Interoperability and Flexibility}
Software citation methods should be sufficiently flexible to accommodate the variant practices among communities, but should not differ so much that they compromise interoperability of software citation practices across communities.
\end{enumerate}
